Title

### 2022-04-27 thesis/experiments/experiments.tex
carlos revision 1

If the reduction clause replaces {\tt \#pragma~omp~crtical}, the second region and the entire program become more efficient, as shown in \cref{fig:pv_regionscomparison_2}. The PaScal Viewer diagrams mentioned in this work, the x-axis (i1--i7) corresponds to different data inputs and the y-axis (2--64) to the number of threads used in the processing. Figures~\ref{fig:pv_regionscomparison} and \ref{fig:pv_regionscomparison_2} were rendered using a tool feature that smoothes the color transition of diagrams. This feature uses interpolation to create a visual element where the color transition is less pronounced. The diagram axes only show the initial and final values with this option. The user can visualize diagrams with only discrete values or even compare the two presentation modes side by side. \cref{fig:visualizationmodes} shows the difference between discrete and smoothed modes considering the whole program and the simulation scenario that uses the clause {\tt \#pragma~omp~crtical}.
\begin{figure}[H]
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_r1.pdf}
		\caption{\centering}
		\label{fig:pv_regionscomparison_a}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_r2.pdf}
		\caption{\centering}
		\label{fig:pv_regionscomparison_b}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_whole.pdf}
		\caption{\centering}
		\label{fig:pv_regionscomparison_c}
	\end{subfigure}
	\caption{Efficiency diagrams and impact of inner regions on program scalability. (x-axis = inputs; y-axis = number of threads). (\textbf{a}) First region diagram. (\textbf{b}) Second region diagram. (\textbf{c}) Whole program~diagram.}
	\label{fig:pv_regionscomparison}
\end{figure}
\unskip
\begin{figure}[H]
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_1.pdf}
		\caption{\centering}
		\label{fig:pv_regionscomparison_a_2}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_2.pdf}
		\caption{\centering}
		\label{fig:pv_regionscomparison_b_2}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_3.pdf}
		\caption{\centering}
		\label{fig:pv_regionscomparison_c_2}
	\end{subfigure}
	\caption{Efficiency diagrams after removing {\tt \#pragma~omp~crtical} clause. (x-axis = inputs; y-axis = number of threads). (\textbf{a}) First region diagram. (\textbf{b}) Second region diagram. (\textbf{c}) Whole program~diagram.}
	\label{fig:pv_regionscomparison_2}
\end{figure}
\unskip
\begin{figure}[H]
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/diagram_discretevalues.pdf}
		\caption{\centering}
		\label{fig:discretevalues}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/diagram_smoothedvalues.pdf}
		\caption{\centering}
		\label{fig:smoothedvalues}
	\end{subfigure}

@@

\textls[-19]{Other analysis objectives not supported by PaScal Viewer, such as visualization of the speedup curve or energy consumption, can be supported by traditional plots. Figures~\ref{fig:speedup} and \ref{fig:energy}} present the charts rendered for analysis by Raytrace and OpenMC applications. In these figures, it is possible to observe that the program efficiency varies according to the increase in the number of threads and the execution of inputs with higher processing loads. From the diagrams in \cref{fig:efficiency_all}, it is possible to observe that the applications present different behaviors concerning their efficiency variations and their scalability capabilities. OpenMC maintains its efficiency almost constant. This pattern represents strong scalability and indicates that the program can maintain its efficiency level when it uses a larger number of threads and processing larger inputs. On the other hand, \cref{fig:efficiency_all}a demonstrates that Raytrace achieves higher efficiency values when processing larger inputs. However, it is also possible to observe that increasing the number of processing units fixing the input size will not improve or hold the efficiency value.
\begin{figure}[H]
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/speedup_completo_rtview_2 (1).pdf}
		\caption{\centering}
		\label{fig:speedup_raytrace}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/speedup_completo_openmc_kernel_novo (1).pdf}
		\caption{\centering}
		\label{fig:speedup_openmc}
	\end{subfigure}
	\caption{\textls[-30]{Speedup of the applications for several input sizes. (\textbf{a}) Raytrace speedup. (\textbf{b}) OpenMC speedup. }}
	\label{fig:speedup}
\end{figure}
\unskip
\begin{figure}[H]
	\begin{subfigure}[b]{0.46\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/energy_completo_rtview_2 (1).pdf}
		\caption{\centering}
		\label{fig:raytrace_en}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.46\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/energy_completo_openmc_kernel_novo (1).pdf}
		\caption{\centering}
		\label{fig:openmc_en}
	\end{subfigure}

	\caption{Energy consumption identified in the execution of applications while varying the number of cores for several input sizes.
		%MDPI: Please add commas to numbers to indicate thousand on the Oy axis of the figures, e.g. 20,000 30,000 40,000.
		(\textbf{a}) Raytrace energy consumption. (\textbf{b}) OpenMC energy consumption.}
	\label{fig:energy}
\end{figure}
\unskip
%Furthermore, using an external graphical interface module, we can load the .json file and intuitively visualize the program's scalability showing the efficiency. The efficiency pinpoint how efficiently a program can take advantage of increasing processing elements on a parallel system. It is defined as the ratio of speedup to the number of processors.
%For example, Figure \ref{fig:efficiency_all} shows examples of graphics generated for the same applications as above. In these figures, it is possible to observe that the program efficiency varies according to the increase in the number of threads and the execution of inputs with higher processing loads. From the Figure \ref{fig:efficiency_raytrace} we can see that as we increase the input size, the efficiency increases while slowly decreasing when the number of cores grows, indicating a tendency for a scalable program. In the second Figure \ref{fig:efficiency_openmc} the efficiency is practically constant while varying the number of cores and the input indicating a strong scalable program.
\begin{figure}[H]
	\begin{subfigure}[b]{0.46\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_raytrace.png}
		\caption{\centering}
		\label{fig:efficiency_raytrace}
	\end{subfigure}
	\begin{subfigure}[b]{0.46\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_openmc.png}
		\caption{\centering}
		\label{fig:efficiency_openmc}
	\end{subfigure}

@@

\section{Validation the model} \label{sec:experimentalvalidation}
In this section, the models presented in~Sections \ref{sec:power_model} and~\ref{sec:performance_model} were validated with a benchmark specific for multi-core architectures. Additionally, in order to assess the modeling overhead and accuracy, our proposal was then compared to machine learning approaches. We compared against support vector regression (SVR)~\cite{Smola2004ARegression}, decision tree \cite{Kitts2006RegressionLecture}, k-nearest neighbors \cite{Altman1992AnRegression}, multilayer perceptron \cite{Murtagh1991MultilayerRegression}, and some new methods, such as Gao et al. \cite{Gao2019DendriticPrediction}. However, SVR was chosen as the most representative because it performed best in our tests without aggressive fine-tuning, as shown in \cref{fig:ml_models}.

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{experiments/figures/ml_models.pdf}
	\caption{Average of the mean squared error for all applications of our study case Section \ref{sec:casestudyapplication}.}
	\label{fig:ml_models}
\end{figure}

\subsection{Verifying Hypothesis}

In this section, we validate whether the assumptions of our model are valid for the system used.

\subsubsection{Frequency and Voltage Relation}
One of the assumptions was that the frequency and the voltage have a linear relationship, as indicated by ~\cref{eq:f_v}. To verify that, we build an experiment that sets the frequency to a specific value while sampling the voltage using the APERF and MPERF registers that provide feedback on the current CPU frequency. The average result of the sampling voltages is shown in  \cref{fig:freq_volt_rel}, where we can observe a near-perfect linear relation. This is because manufacturers implement this curve in the processors, using tables that relate ranges of frequencies to voltages so that they can precisely define any curve that will better suit their design.

\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}
	\includegraphics[width=\columnwidth]{experiments/figures/freq_volt_rel.pdf}
	\caption{Frequency voltage relation.}
	\label{fig:freq_volt_rel}
\end{figure}

\subsubsection{Input Size and Instructions}
We ran the applications with different inputs assuming  linear growth in the amount of work for one input to the other when building our model. However, measuring and controlling the amount of work would require much instrumentation and tuning to find an input corresponding to a certain amount of work. Therefore, to build our models, we use the time to reference the amount of work, assuming that the work is proportional to the executing time. \cref{fig:time_input} corresponds to the verification of this supposition.

\cref{tab:corr_in_time} shows that the assumption was reasonable since the average correlation was 0.96 for all applications, indicating that growth in the number of instructions  will follow the time. This was the case for all applications that we ran in our benchmark and should hold for any data parallelism type of application.
\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/blackscholes.pdf}
		\caption{Blackscholes.}
		\label{fig:time_input_blackshoels}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/canneal.pdf}
		\caption{Canneal.}
		\label{fig:time_input_canneal}
	\end{subfigure}

	\hfill
	\caption{Relation between time and instructions for each input size. }
	\label{fig:time_input}
\end{figure}
\vspace{-6pt}

\begin{table}[H]
	\centering
	\caption{Correlation of time and instructions for all applications.}
	\begin{tabular}{ll}
		\multicolumn{1}{l|}{\textbf{Application}} & \textbf{Correlation} \\ \hline
		Blackscholes                                               & 0.99                                  \\
		Bodytrack                                                  & 0.99                                  \\
		Canneal                                                    & 0.99                                  \\
		Dedup                                                      & 0.99                                  \\
		Ferret                                                     & 0.96                                  \\
		Fluidanimate                                               & 0.99                                  \\
		Freqmineq                                                  & 0.99                                  \\
		Openmc                                                     & 0.94                                  \\
		Raytrace                                                   & 0.99                                  \\
		Swaptions                                                  & 0.99                                  \\
		Vips                                                       & 0.98                                  \\
		x264                                                       & 0.99                                  \\
		HPL                                                        & 0.79                                  \\ \hline
	\end{tabular}
	\label{tab:corr_in_time}
\end{table}

The next assumption was that the application's behavior was the same when varying the workload. This condition is necessary for using the model with an unknown input size because, if the behavior is the same, we can interpolate the known inputs. One way to verify this is to measure the rate of instructions per second normalized by the frequency, as shown in \cref{fig:fingerpritns}.
\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}

	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/blackscholes.pdf}
		\caption{Blacksholes.}
		\label{fig:fp_baclscholes}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/canneal.pdf}
		\caption{Canneal.}
		\label{fig:fp_canneal}
	\end{subfigure}

	\hfill
	\caption{Rate of instructions per second varying the input size normalized by the frequency.}
	\label{fig:fingerpritns}
\end{figure}

\cref{fig:fingerpritns} shows that the applications have roughly the same curve when normalized; this also happens for all other applications in our benchmark.

The final assumption is that the workload should also not vary depending on the number of cores or frequency. To verify, we measure the total number of executed instructions while varying the cores from 1 to 32. \cref{tab:cores_variation} shows the results.

\begin{table}[H]
	\centering
	\caption{Variation of the number of instructions when changing the number of cores for the same input.}
	%\begin{tabular}{c|c|c|c}
	\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}cccc}
		\toprule
		& \textbf{Average} & \textbf{Standard}     & \textbf{Standard}\\
		\multirow{-2}{*}{\textbf{Application}}& \textbf{Number of Instructions} & \textbf{Deviation}     & \textbf{Deviation} \textbf{(\%) }\\ \midrule
		Vip          & $7.97 \times 10^{11}$     & $7.16 \times 10^{6}$  & 0.00 \\ %Please use Scientific notation. For example, “1.65e+04” should be changed to “1.65× 104”. Please confirm `%` whether to remove.
		Openmc       & $8.17 \times 10^{7}$     & $ 1.65 \times 10^{4}$ & 0.02    \\ 
		Rtview       & $9.91 \times 10^{12}$     & $ 1.55 \times 10^{9}$ & 0.02    \\ 
		X264         & $4.52 \times 10^{11}$     & $ 5.81 \times 10^{7}$ & 0.01    \\ 
		Bodytrack    & $1.86 \times 10^{12}$     & $ 3.95 \times 10^{10}$ & 2.13    \\ 
		Fluidanimate & $2.09 \times 10^{12}$     & $ 8.44 \times 10^{10}$ & 4.04    \\ 
		HPL          & $1.14 \times 10^{8}$      & $ 1.24 \times 10^{5}$ & 0.11    \\ 
		Blackschole  & $3.75 \times 10^{12}$     & $ 1.40 \times 10^{9}$ & 0.04     \\ 
		Dedup        & $1.02 \times 10^{11}$     & $ 5.74 \times 10^{7}$ & 0.06     \\ 
		Swapti       & $2.43 \times 10^{12}$     & $ 8.87 \times 10^{8}$ & 0.04     \\ 
		Canneal      & $1.19 \times 10^{11}$     & $ 4.46 \times 10^{7}$ & 0.04     \\ 
		Freqmine     & $1.27 \times 10^{12}$     & $ 4.78 \times 10^{8}$ & 0.04     \\ 
		Ferret       & $4.76 \times 10^{11}$     & $ 7.04 \times 10^{7}$ & 0.01     \\ \bottomrule
	\end{tabular*}
	\label{tab:cores_variation}
\end{table}
% Corrected

\cref{tab:cores_variation} shows the standard deviation and what that corresponds to in terms of the total number of instructions as a percentage.

The same test was performed for the frequency, varying from 1.2 to 2.2 GHz with 100~MHz steps. The results are shown in \cref{tab:freq_variation}.

These results show that all the assumptions were reasonable, and we can safely move to the validation of the model's prediction.
\begin{table}[H]
	\caption{Variation of the number of instructions when changing the frequency for the same input.}
	\centering
	%\begin{tabular}{c|c|c|c}{\hsize}{@{\extracolsep{\fill}}cccc}
	\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}cccc}
		\toprule
		& \textbf{Average} & \textbf{Standard}     & \textbf{Standard}\\
		\multirow{-2}{*}{\textbf{Application}}& \textbf{Number of Instructions} & \textbf{Deviation}     & \textbf{Deviation} \textbf{(\%)} \\ \midrule
		Vip          & $7.97 \times 10^{11}$      & $ 1.16 \times 10^{6}$  & 0.00      \\ 
		Openmc       & $8.17 \times 10^{7}$       & $ 4.52 \times 10^{3}$  & 0.01      \\ 
		Rtview       & $9.91 \times 10^{12}$      & $ 6.64 \times 10^{5}$  & 0.00      \\ 
		X264         & $4.52 \times 10^{11}$      & $ 1.54 \times 10^{5}$  & 0.00      \\ 
		Bodytrack    & $1.84 \times 10^{12}$      & $ 2.54 \times 10^{5}$  & 0.00      \\ 
		Fluidanimate & $2.38 \times 10^{12}$      & $ 1.70 \times 10^{9}$  & 0.07      \\ 
		HPL          & $1.14 \times 10^{8}$       & $ 5.95 \times 10^{3}$  & 0.01      \\ 
		Blackschole  & $3.75 \times 10^{12}$      & $ 4.36 \times 10^{5}$  & 0.00      \\ 
		Dedup        & $1.02 \times 10^{11}$      & $ 8.32 \times 10^{7}$  & 0.08      \\ 
		Swapti       & $2.43 \times 10^{12}$      & $ 1.48 \times 10^{5}$  & 0.00      \\ 
		Canneal      & $1.19 \times 10^{11}$      & $ 3.01 \times 10^{5}$  & 0.00      \\ 
		Freqmine     & $1.27 \times 10^{12}$      & $ 3.70 \times 10^{8}$  & 0.03      \\ 
		Ferret       & $4.76 \times 10^{11}$      & $ 5.63 \times 10^{7}$  & 0.01      \\\bottomrule
	\end{tabular*}
	\label{tab:freq_variation}
\end{table}

\subsection{Fitting the Models} \label{subsec:fitting}
To find the parameters of  \cref{eq:en_final}, 10 uniformly random configurations of frequencies ($f$), cores ($p$) and inputs ($N$) were chosen from the range $1<=p<=32$, $1.2<=f<=2.2$ and $1<=N<=5$, respectively. The application was executed for each chosen configuration, and the measured energy and time values were collected. For the input size, if we assume that all CPU instructions take approximately the same time to execute, the number of basic operations will be directly correlated with the time. Thus, we can estimate the input size by looking at the execution time, allowing us to divide a large input size into several smaller ones, knowing their relationship, as performed in the work of Oliveira ~\cite{Oliveira2018ApplicationCharacterization}. The unity can also vary depending on the definition. For simplicity, we assign numbers from 1 to 10, increasing the problem linearly, so it is also possible to interpolate any input in between these values.

For each configuration, samples of the power were collected using IPMI every 1 second. This sampling rate was chosen based on the magnitude of the mean run time of the applications, which is in the order of minutes. Therefore, this rate provides enough samples to measure average power. Additionally, timestamps and the total run time were collected. The total energy spent on each configuration is estimated by first interpolating the power samples using the first-order method and then integrating this function in the time.

The model's parameters are calculated by solving an optimization problem of finding the values that minimize the squared error of the prediction to the measured values using the non-linear least-squares method.

The Python library Scikit-Learn was used to build the SVR model~\cite{Pedregosa2011Scikit-learn:Python}. The SVR was trained using the same data used for parameter estimation of Equation (\ref{eq:en_final}) %\cref{eq:en_final}
with a grid search used to find the best kernel function and the best values for the hyper-parameters penalty for the wrong ($C$) and ($\gamma$). For this data, the best function was the radial base function (RBF), and the hyper-parameters were $C=10^4$ and $\gamma=0.5$.

\subsection{Measured versus Modeled Energy}
\label{subsec:measuredversusmodeledenergy}

To validate the model, we ran all possible configurations in the tested machine, varying the cores in a range of $1<=p<=32$, the frequency in $1.2<=f<=2.2$, and the input in $1<=N<=5$. The total number of configurations varies from 400 to over 1000 depending on the application, as some applications have restrictions on the number of cores that they can run. Once the data was collected, we computed the mean percentage error (MPE) according to the following equation:
\begin{equation}
	MPE = \frac{1}{N} \sum_i^N \frac{|y_{\rm estimated}-y_{\rm measured}|}{y_{\rm measured}}.
	\label{eq:mpe}
\end{equation}

\subsubsection{Frequency $\times$ Cores}
\cref{fig:en_eq_freq_cores_fc} plots the measured and modeled energy consumption for some of the applications modeled. In addition,  some of the possible shapes that the model can take while varying the number of active cores, and operating frequency, are shown.%please confirm that the intended meaning has been retained.
% yes its the same

\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_cores/completo_black_5.pdf}}
		\caption{}
		\label{fig:en_eq_black_fc}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_cores/completo_canneal_1.pdf}}
		\caption{}
		\label{fig:en_eq_canneal_fc}
	\end{subfigure}

	\caption{Example fit for a specific input size: Blackscholes (\textbf{a}) and Canneal (\textbf{b}).  “measured values” are the sensor data, and “minimum energy” is the minimum energy model prediction.
	}
	\label{fig:en_eq_freq_cores_fc}
\end{figure}
\subsubsection{Frequency $\times$ Input}
\cref{fig:en_eq_freq_inp_fi} plots the measured and modeled energy consumption for some of the applications modeled. The diagrams show some of the possible shapes that the model can take while varying the operating frequency, and input size.
\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_inps/completo_black_5.pdf}}
		\caption{}
		\label{fig:en_eq_black_fi}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_inps/completo_canneal_1.pdf}}
		\caption{}
		\label{fig:en_eq_canneal_fi}
	\end{subfigure}

	\caption{Example fit for a specific input size: Blackscholes (\textbf{a}) and Canneal (\textbf{b}).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
	}
	\label{fig:en_eq_freq_inp_fi}
\end{figure}
\subsubsection{Cores $\times$ Input}
\cref{fig:en_eq_core_inp_ci} plots the measured and modeled energy consumption for some of the applications modeled. The diagrams  show some of the possible shapes that the model can take while varying the number of active cores, and input size.
\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/cores_inps/completo_black_5.pdf}}
		\caption{}
		\label{fig:en_eq_black_ci}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/cores_inps/completo_canneal_1.pdf}}
		\caption{}
		\label{fig:en_eq_canneal_ci}
	\end{subfigure}

	\caption{Example fit for a specific input size: Blackscholes (\textbf{a}) and Canneal (\textbf{b}).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
	}
	\label{fig:en_eq_core_inp_ci}
\end{figure}

\subsubsection{Comparison}
The average results for each application were calculated using a model trained with only 10 configurations, and the comparison is displayed \cref{fig:mpe_svr_eq}. %We calculated the raw MPE values shown in ~\cref{tab:mpe_svr_eq}.
\begin{figure}[H]
	\centering
	\includegraphics[width=.8\columnwidth]{models/figures/mpe_svr_eq.pdf}
	\caption{Comparison of the mean percentage error between the proposed model and SVR. ``Model mean'' and ``SVR mean'' are the average of all MPE values for all applications.
	}
	\label{fig:mpe_svr_eq}
\end{figure}

\cref{fig:mpe_svr_eq} shows that the proposed model always performed better, with a lower MPE than SVR, when we were limited to 10 training points. This result is further explored in the next Section \ref{subsec:overhead}, where we undertake a comparison with different training sizes. The exact values are shown in \cref{tab:mpe_svr_eq}.

\begin{table}[H]
\centering
\begin{tabular}{c|c|c}
\hline
Application  & Model & SVR   \\ \hline
Ferret       & 5.25     & 12.49  \\ 
Raytrace     & 6.36     & 11.95  \\
Fluianimate  & 2.44     & 22.90  \\
x264         & 8.28     & 15.33  \\
Vips         & 7.54     & 10.80  \\
Swaptions    & 6.54     & 18.57  \\
Canneal      & 3.12     & 6.13   \\
Dedup        & 8.85     & 13.70  \\
Freqmine     & 2.44     & 3.24   \\
Blackscholes & 2.18     & 11.00  \\
HPL          & 7.47     & 12.75  \\
Bodytrack    & 16.98    & 34.12  \\
Openmc       & 11.15    & 24.34  \\
\end{tabular}
\caption{Comparison of the Mean Percentage Error between the proposed model and SVR: raw values.}
\label{tab:mpe_svr_eq}
\end{table}

\subsection{Overheads on Training} \label{subsec:overhead}
It is known that machine learning is data-driven; in that sense, the SVR model obtained using only 10 configurations could be improved, but what about the analytical model? 
To answer that question, the proposed model and the SVR were also trained with a varying number of configurations.
We then compared the MPE and the amount of energy spent to create each model. 
This accuracy-energy trade-off is crucial since building models' energy overhead defeats the primary goal of saving power when running applications.
\enlargethispage{.5cm}

\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_ferret_4.pdf}}
		\caption{MPE for Ferret.}
		\label{fig:overhead_ferret}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_black_6.pdf}}
		\caption{MPE for Blackscholes.}
		\label{fig:overhead_black}
	\end{subfigure}
	%
	\par\bigskip
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_x264_4.pdf}}
		\caption{MPE for x264.}
		\label{fig:overhead_x264}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_dedup_4.pdf}}
		\caption{MPE for Dedup.}
		\label{fig:overhead_dedup}
	\end{subfigure}

	\caption{MPE of the case studies versus training size, comparing how many training points is necessary to reach an acceptable result.}
	\label{fig:overheadapps}
\end{figure}

>>>>

If the reduction clause replaces {\tt \#pragma~omp~crtical}, the second region and the entire program become more efficient, as shown in \cref{fig:pv_regionscomparison_2}. The PaScal Viewer diagrams mentioned in this work, the x-axis (i1--i7) corresponds to different data inputs and the y-axis (2--64) to the number of threads used in the processing. Figures~\ref{fig:pv_regionscomparison} and \ref{fig:pv_regionscomparison_2} were rendered using a tool feature that smoothes the color transition of diagrams. This feature uses interpolation to create a visual element where the color transition is less pronounced. The diagram axes only show the initial and final values with this option. The user can visualize diagrams with only discrete values or even compare the two presentation modes side by side. \cref{fig:visualizationmodes} shows the difference between discrete and smoothed modes considering the whole program and the simulation scenario that uses the clause {\tt \#pragma~omp~crtical}.
\begin{figure}[H]
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_r1.pdf}
%		\caption{\centering}
		\label{fig:pv_regionscomparison_a}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_r2.pdf}
%		\caption{\centering}
		\label{fig:pv_regionscomparison_b}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_whole.pdf}
%		\caption{\centering}
		\label{fig:pv_regionscomparison_c}
	\end{subfigure}
	\caption{Efficiency diagrams and impact of inner regions on program scalability. (x-axis = inputs; y-axis = number of threads). (\textbf{a}) First region diagram. (\textbf{b}) Second region diagram. (\textbf{c}) Whole program~diagram.}
	\label{fig:pv_regionscomparison}
\end{figure}
\unskip
\begin{figure}[H]
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_1.pdf}
%		\caption{\centering}
		\label{fig:pv_regionscomparison_a_2}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_2.pdf}
%		\caption{\centering}
		\label{fig:pv_regionscomparison_b_2}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_3.pdf}
%		\caption{\centering}
		\label{fig:pv_regionscomparison_c_2}
	\end{subfigure}
	\caption{Efficiency diagrams after removing {\tt \#pragma~omp~crtical} clause. (x-axis = inputs; y-axis = number of threads). (\textbf{a}) First region diagram. (\textbf{b}) Second region diagram. (\textbf{c}) Whole program~diagram.}
	\label{fig:pv_regionscomparison_2}
\end{figure}
\unskip
\begin{figure}[H]
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/diagram_discretevalues.pdf}
%		\caption{\centering}
		\label{fig:discretevalues}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/diagram_smoothedvalues.pdf}
%		\caption{\centering}
		\label{fig:smoothedvalues}
	\end{subfigure}

@@

\textls[-19]{Other analysis objectives not supported by PaScal Viewer, such as visualization of the speedup curve or energy consumption, can be supported by traditional plots. Figures~\ref{fig:speedup} and \ref{fig:energy}} present the charts rendered for analysis by Raytrace and OpenMC applications. In these figures, it is possible to observe that the program efficiency varies according to the increase in the number of threads and the execution of inputs with higher processing loads. From the diagrams in \cref{fig:efficiency_all}, it is possible to observe that the applications present different behaviors concerning their efficiency variations and their scalability capabilities. OpenMC maintains its efficiency almost constant. This pattern represents strong scalability and indicates that the program can maintain its efficiency level when it uses a larger number of threads and processing larger inputs. On the other hand, \cref{fig:efficiency_all}a demonstrates that Raytrace achieves higher efficiency values when processing larger inputs. However, it is also possible to observe that increasing the number of processing units fixing the input size will not improve or hold the efficiency value.
\begin{figure}[H]
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/speedup_completo_rtview_2 (1).pdf}
%		\caption{\centering}
		\label{fig:speedup_raytrace}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/speedup_completo_openmc_kernel_novo (1).pdf}
%		\caption{\centering}
		\label{fig:speedup_openmc}
	\end{subfigure}
	\caption{\textls[-30]{Speedup of the applications for several input sizes. (\textbf{a}) Raytrace speedup. (\textbf{b}) OpenMC speedup. }}
	\label{fig:speedup}
\end{figure}
\unskip
\begin{figure}[H]
	\begin{subfigure}[b]{0.46\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/energy_completo_rtview_2 (1).pdf}
%		\caption{\centering}
		\label{fig:raytrace_en}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.46\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/energy_completo_openmc_kernel_novo (1).pdf}
%		\caption{\centering}
		\label{fig:openmc_en}
	\end{subfigure}

	\caption{Energy consumption identified in the execution of applications while varying the number of cores for several input sizes.
		%MDPI: Please add commas to numbers to indicate thousand on the Oy axis of the figures, e.g. 20,000 30,000 40,000.
		(\textbf{a}) Raytrace energy consumption. (\textbf{b}) OpenMC energy consumption.}
	\label{fig:energy}
\end{figure}
\unskip
%Furthermore, using an external graphical interface module, we can load the .json file and intuitively visualize the program's scalability showing the efficiency. The efficiency pinpoint how efficiently a program can take advantage of increasing processing elements on a parallel system. It is defined as the ratio of speedup to the number of processors.
%For example, Figure \ref{fig:efficiency_all} shows examples of graphics generated for the same applications as above. In these figures, it is possible to observe that the program efficiency varies according to the increase in the number of threads and the execution of inputs with higher processing loads. From the Figure \ref{fig:efficiency_raytrace} we can see that as we increase the input size, the efficiency increases while slowly decreasing when the number of cores grows, indicating a tendency for a scalable program. In the second Figure \ref{fig:efficiency_openmc} the efficiency is practically constant while varying the number of cores and the input indicating a strong scalable program.
\begin{figure}[H]
	\begin{subfigure}[b]{0.46\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_raytrace.png}
%		\caption{\centering}
		\label{fig:efficiency_raytrace}
	\end{subfigure}
	\begin{subfigure}[b]{0.46\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_openmc.png}
%		\caption{\centering}
		\label{fig:efficiency_openmc}
	\end{subfigure}

@@
